---
title: "Data Package Prep"
output: html_document
date: "2023-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE)

library(tidyverse)
library(plotly)
```

---

### FOLLOW-UP ACTION ITEMS:

1. soil-ghg and sediment-ghg: Site -> Transect; type -> treatment

---


```{r import}

import_files = function(FILEPATH){
  
  filePaths <- list.files(path = FILEPATH, pattern = ".csv", full.names = TRUE)
  
    do.call(bind_rows, lapply(filePaths, function(path){
      # then add a new column `source` to denote the file name
      df <- 
        read.csv(path, header = TRUE, check.names = FALSE) %>% 
        mutate(source = basename(path))
      df}))
  
}

kit_metadata = read.csv("Data/ESS-DIVE Submission/ec1-metadata/ec1_metadata_kitlevel.csv") %>% dplyr::select(kit_id, site_name, region)
reorder_transect = function(dat){
  dat %>% 
    mutate(transect_location = factor(transect_location, levels = c("upland", "transition", "wetland")))
}


soil_combined = import_files(FILEPATH = "Data/ESS-DIVE Submission/ec1-soil")
soil_combined_long = 
  soil_combined %>% 
  pivot_longer(-c(campaign, kit_id, transect_location, source, site, type)) %>% 
  mutate(transect_location = case_when(is.na(transect_location) ~ site, TRUE ~ transect_location),
         transect_location = tolower(transect_location)) %>% 
  dplyr::select(-site) %>% 
  rename(treatment = type) %>% 
  filter(!is.na(value)) %>% 
  left_join(kit_metadata) %>% 
  reorder_transect()

sediment_combined = import_files(FILEPATH = "Data/ESS-DIVE Submission/ec1-sediment")
sediment_combined_long = 
  sediment_combined %>% 
  pivot_longer(-c(campaign, kit_id, transect_location, source, site, type)) %>% 
  mutate(transect_location = case_when(is.na(transect_location) ~ site, TRUE ~ transect_location),
         transect_location = tolower(transect_location)) %>% 
  dplyr::select(-site) %>% 
  rename(treatment = type) %>% 
  filter(!is.na(value)) %>% 
  left_join(kit_metadata) %>% 
  mutate(transect_location = "sediment")


water_combined = import_files(FILEPATH = "Data/ESS-DIVE Submission/ec1-water")
water_combined_long = 
  water_combined %>% 
  pivot_longer(-c(campaign, kit_id, transect_location, source)) %>% 
  mutate(transect_location = tolower(transect_location)) %>% 
  filter(!is.na(value)) %>% 
  left_join(kit_metadata) %>% 
  mutate(transect_location = "water")


```

```{r summary_fns}
plot_data = function(data, TITLE){
  p = 
    data %>% 
    ggplot(aes(x = region, y = value, color = transect_location,
               text = paste("Kit:", kit_id, 
                            "\nSite:", site_name, 
                            "\nValue:", round(value, 2)),
               group = transect_location))+
    geom_point(position = position_dodge(width = 0.3))+
    facet_wrap(~source + name, scales = "free_y")+
    labs(title = TITLE)
  
    ggplotly(p, tooltip = "text")
  
}

summarize_data = function(data){
  data %>% 
    group_by(source, name, region) %>% 
    dplyr::summarise(mean = round(mean(value),3),
                     median = round(median(value),3),
                     min = round(min(value),3),
                     max = round(max(value),3),
                     n = n()) 
  
}

```

---

### Water

```{r water, fig.height=7, fig.width=11}

plot_data(data = water_combined_long, TITLE = "Water")

summarize_data(data = water_combined_long) %>% knitr::kable()
```


---

### Sediment

```{r sediment, fig.height=7, fig.width=11}

plot_data(data = sediment_combined_long, TITLE = "Sediment")

summarize_data(data = sediment_combined_long) %>% knitr::kable()
```


---

### Soil

```{r soil, fig.height=7, fig.width=11}

plot_data(data = soil_combined_long, TITLE = "Soil")

summarize_data(data = soil_combined_long) %>% knitr::kable()
```

---

---


2023-01-17, KFP comments:

**soil**

- BD: 
  - K049 upland (Saugatuch) is high for CB, at 1.74 
- GHG/pCH4: 
  - K050 wetland (Kimages) is high, at 20.68
  - K015 upland (Taskinas) is high, at 13.68
- GHG/pCO2:
  - K015 upland (Taskinas) is high, at 208251
- GHG/pN2O
  - K018 (Goodwin) and K034 (King's Creek) uplands are high, at > 500
  - K019 (Boy Scout) upland is high, at 590
- LOI: there is a group in CB that is higher than the rest: 
  - K039 T and W, 
  - K047 W 
  - K008 T
- O2 drawdown:
  - K059 transition (Blackwater) is higher than other transitions
- pH
  - K013 upland (Wigwam Bay) is higher than the rest, at 7.56
  
**sediment**

- GHG/pCH4: K021 (Magee Marsh) is way higher than the rest, at 114
- GHG/pN2O: K021 (Magee Marsh) is way higher than the other CB, at 16.49
- O2 drawdown: very high spread, with K030 (St. Jones Reserve) the highest, at 48



---

---

## Session Info 

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>