---
title: "EXCHANGE Campign 1: Exploratory Figures"
output: 
  github_document:
      html_preview: true
  

---

Date updated: `r Sys.Date()`


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      collapse = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      comment = "#>")
```

```{r packages}
# load packages 
library(tidyverse)
theme_set(theme_bw() + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12), axis.title.x = element_blank()))
```

```{r metadata}
# import kit-level metadata

metadata_kit_temp = read.csv("Data/EC1_Metadata_KitLevel.csv")
metadata_kit = 
  metadata_kit_temp %>% 
  dplyr::select(kit_id, site_name, region) %>% 
  mutate(kit_id = str_pad(kit_id, 3, pad = "0"),
         kit_id = paste0("K", kit_id)) %>% 
  #  mutate(region = recode(region, "Chesapeake Bay" = "CB", "Great Lakes" = "GL")) %>% 
  force()
```


```{r import_data}
# import data
import_data = function(FILEPATH){
  
  # pull a list of file names in the target folder with the target pattern
  # then read all files and combine
  
  filePaths <- list.files(path = FILEPATH, pattern = ".csv", full.names = TRUE)
  
  # dat <- 
  do.call(bind_rows, lapply(filePaths, function(path){
    # then add a new column `source` to denote the file name
    df <- read.csv(path, header = TRUE)
    #  df <- read.delim(path, skip = 2)
    df[["source"]] <- rep(path, nrow(df))
    df}))
  
}

# Now, run this function
all_data <- import_data(FILEPATH = "Data/Processed/")

all_data_processed = 
  all_data %>% 
  mutate(transect_location = tolower(transect_location)) %>% 
  dplyr::select(#-ends_with("flag"),
                -starts_with("date")) %>% 
  separate(source, sep = "EC1_", into = c("source1", "analysis")) %>% 
  mutate(analysis = str_remove(analysis, "_[0-9]{8}"),
         analysis = str_remove(analysis, ".csv"),
         analysis = str_remove(analysis, "_L[0-9][A-Z]")) %>% 
  dplyr::select(-source1) %>% 
  mutate(analysis = str_remove(analysis, "Soil_"),
         analysis = str_remove(analysis, "Water_")) %>% 
  mutate(transect_location = factor(transect_location, levels = c("upland", "transition", "wetland", "sediment", "water"))) %>% 
  arrange(kit_id, transect_location) %>% 
  filter(!is.na(kit_id)) %>% 
  left_join(metadata_kit) %>% 
  filter(!is.na(region))

blank_kit =
  all_data_processed %>% 
  filter(is.na(region)) %>% 
  distinct(kit_id)

```


```{r map_function}
## This script is designed to make several figures to be used for looking at the
## data coming from EC1 with collaborators at the 2.2 Workshop in 02/22
##
## 2022-02-15 (updated 6/2/22 to be project-agnostic)
## Peter Regier 
## 
# #############
# #############

# 1. Setup ---------------------------------------------------------------------

## Load packages
require(pacman)
p_load(cowplot, ## plot_grid()
       tidyverse, 
       sf, ## simple features
       janitor, # clean_names()
       stringr,
       googlesheets4, 
       #googledrive, ## import data from Google Sheets
       ggthemes) ## theme_map()


# load lat-lon file
latlong <- read_sheet("https://docs.google.com/spreadsheets/d/1_tKn5vHryVMHvuQix_PRXhzQcEkeAG5LDCBXjs7nZEc/edit#gid=842964887", sheet = 1) %>% 
  mutate(kit_id = paste0("K", Kit_ID)) %>% 
  rename("lat" = `Water_Latitude`, 
         "long" = `Water_Longitude`) %>% 
  select(kit_id, lat, long)


make_map = function(DAT, VAR, COLOR_LEGEND){
  
  ## Set CRS
  common_crs <- 4326
  
  ## Set map size and point size
  point_size <- 2
  map_width = 9
  map_height = 6

  # Set up map layers for plotting ------------------------------------------
  
  ## Set regional and CB (inset) bounding boxes
  us_bbox <- c(xmin = -125, xmax = -60, ymin = 20, ymax = 50)
  region_bbox <- c(xmin = -95, xmax = -70, ymin = 35, ymax = 48)
  cb_bbox <- c(xmin = -77.8, xmax = -74.5, ymin = 36.8, ymax = 40.5)
  
  ## Make US states map cropped to GL/CB region
  us <- read_sf("Reports_markdown/cb_2018_us_state_5m/cb_2018_us_state_5m.shp") %>% 
    st_transform(., crs = common_crs) %>% 
    st_crop(., y = us_bbox)
  
  region <- st_crop(us, y = region_bbox)
  
  ## Further crop states to CB region
  cb_states <- st_crop(region, y = cb_bbox)
  
  
  # make a dataset merging metadata with site lat-longs
  df_map <- inner_join(DAT, latlong, by = "kit_id") %>% 
    st_as_sf(., coords = c("long", "lat"), crs = common_crs)
  
  ## Crop data  to CB region for inset plot
  df_cb <- st_crop(df_map, y = cb_bbox)
  
  ## get max and min values for variable
  var_min = DAT %>% dplyr::select(-kit_id) %>% pull() %>% min(na.rm = TRUE)
  var_max = DAT %>% dplyr::select(-kit_id) %>% pull() %>% max(na.rm = TRUE)

  
  ## Make the base map with all sites
  base_plot <- 
    ggplot() + 
    geom_sf(data = region) + 
    #geom_sf(data = meta_map, size = point_size * 3, color = "white") +
    geom_sf(data = df_map, 
            color = "white", size = point_size * 2.5) + 
    geom_sf(data = df_map, 
            aes_string(color = VAR), 
            size = point_size * 2) + 
    geom_rect(aes(xmin = -77.8, xmax = -74.5, ymin = 36.8, ymax = 40.5), 
              fill = NA, color = "black", lwd = 0.75) +
    geom_segment(aes(x = -77.8, xend = -76, y = 40.5, yend = 42), 
                 color = "black", lwd = 0.75) + 
    geom_segment(aes(x = -74.5, xend = -71, y = 40.5, yend = 41.5), 
                 color = "black", lwd = 0.75) + 
    theme_map() + 
    theme(legend.background = element_rect(fill = alpha("white", 0.0)), 
          legend.key = element_rect(fill = "transparent"), 
          legend.position = c(0.85, 0.1)) + 
    scale_color_viridis_c(limits = c(var_min, var_max)) + 
    labs(x = "", y = "" , color = COLOR_LEGEND) + 
    grattantheme::watermark("Preliminary", fontsize = 40)
  
  ## Make the inset map with just CB sites
  inset_plot <- 
    ggplot() + 
    geom_sf(data = cb_states) + 
    geom_sf(data = df_cb, 
            color = "white", size = point_size * 2) + 
    geom_sf(data = df_cb, 
            aes_string(color = VAR), 
            size = point_size * 1.7) + 
    theme_map() + 
    theme(legend.background = element_rect(fill = alpha("white", 0.0)), 
          legend.key = element_rect(fill = "transparent"), 
          #legend.position = c(0.7, 0)
          legend.position = "none") + 
    scale_color_viridis_c(limits = c(var_min, var_max)) + 
    theme(panel.background = element_rect(fill = "white", colour = "black")) + 
    labs(color = "")
  
  ## Combine into single figure
  base_plot + 
    annotation_custom(
      ggplotGrob(inset_plot), 
      xmin = -78, xmax = -70, ymin = 41, ymax = 48.8)
 
  
  
  
     
}

```


---

# Soils and sediments

## Soil: total C and N

```{r tctn}
tctn_data = 
  all_data_processed %>% 
  filter(analysis == "TCTN") %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) 
```

```{r tctn_jitter, fig.width=8, fig.height=4}
tctn_data %>% 
  ggplot(aes(x = transect_location, y = tc_perc))+
  geom_jitter(width = 0.1)+
  facet_wrap(~region)+
  labs(x = "",
       y = "Total C, %")

tctn_data %>% 
  ggplot(aes(x = transect_location, y = tn_perc))+
  geom_jitter(width = 0.1)+
  facet_wrap(~region)+
  labs(x = "",
       y = "Total N, %")

```

```{r tctn_map}
make_map(DAT = tctn_data %>% filter(transect_location == "upland") %>% dplyr::select(kit_id, tc_perc), VAR = "tc_perc", COLOR_LEGEND = "Total C, %") + ggtitle("Upland soils: total C")
make_map(DAT = tctn_data %>% filter(transect_location == "upland") %>% dplyr::select(kit_id, tn_perc), VAR = "tn_perc", COLOR_LEGEND = "Total N, %") + ggtitle("Upland soils: total N")
```


---


## Soil: pH and specific conductance

```{r pH}
pH_data = 
  all_data_processed %>% 
  filter(analysis == "pH") %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) 
```

```{r pH_jitter, fig.width=8, fig.height=4}
pH_data %>% 
  ggplot(aes(x = transect_location, y = ph))+
  geom_jitter(width = 0.1)+
  facet_wrap(~region)+
  labs(x = "",
       y = "pH")

pH_data %>% 
  ggplot(aes(x = transect_location, y = specific_conductance_us_cm/1000))+
  geom_jitter(width = 0.2)+
  labs(x = "",
       y = "Specific conductance, mS/cm")+
  facet_wrap(~region, scales = "free_y")
```

```{r pH_map}
make_map(DAT = pH_data %>% filter(transect_location == "upland") %>% dplyr::select(kit_id, ph), VAR = "ph", COLOR_LEGEND = "pH") + ggtitle("Upland soils: pH")
make_map(DAT = pH_data %>% filter(transect_location == "upland") %>% dplyr::select(kit_id, specific_conductance_us_cm) %>% mutate(specific_conductance_ms_cm = specific_conductance_us_cm/1000), 
         VAR = "specific_conductance_ms_cm", COLOR_LEGEND = "Sp. cond., mS/cm") + ggtitle("Upland soils: Sp conductance")
```

---

## Soil: bulk density

```{r bd}
bd_data = 
  all_data_processed %>% 
  filter(analysis == "BulkDensity") %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) 
```

```{r nd_jitter, fig.width=8, fig.height=4}
bd_data %>% 
  ggplot(aes(x = transect_location, y = bulk_density_g_cm3))+
  geom_jitter(width = 0.1)+
  facet_wrap(~region)+
  labs(x = "",
       y = "Bulk density, g/cm3")
```

```{r bd_map}
make_map(DAT = bd_data %>% filter(transect_location == "upland") %>% dplyr::select(kit_id, bulk_density_g_cm3), VAR = "bulk_density_g_cm3", COLOR_LEGEND = "Bulk density, g/cm3") + ggtitle("Upland soils: bulk density")
```

---

## Soil: gravimetric water content

```{r gwc}
gwc_data = 
  all_data_processed %>% 
  filter(analysis == "GWC") %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) 
```

```{r gwc_jitter, fig.width=8, fig.height=4}
gwc_data %>% 
  ggplot(aes(x = transect_location, y = gwc_perc))+
  geom_jitter(width = 0.1)+
  facet_wrap(~region)+
  labs(x = "",
       y = "Gravimetric water, %")
```

```{r gwc_map}
make_map(DAT = gwc_data %>% filter(transect_location == "upland") %>% dplyr::select(kit_id, gwc_perc), VAR = "gwc_perc", COLOR_LEGEND = "gravimetric water, %") + ggtitle("Upland soils: water content")
```

---

## Soil: loss on ignition

```{r loi}
loi_data = 
  all_data_processed %>% 
  filter(analysis == "LOI") %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) 
```

```{r loi_jitter, fig.width=8, fig.height=4}
loi_data %>% 
  ggplot(aes(x = transect_location, y = loi_perc))+
  geom_jitter(width = 0.1)+
  facet_wrap(~region)+
  labs(x = "",
       y = "Loss on ignition, %")
```

```{r loi_map}
make_map(DAT = loi_data %>% filter(transect_location == "upland") %>% dplyr::select(kit_id, loi_perc), VAR = "loi_perc", COLOR_LEGEND = "loss on ignition, %") + ggtitle("Upland soils: loss on ignition")
```

---

---

# Surface waters

## Water: Ions

```{r ions}
ions_data = 
  all_data_processed %>% 
  filter(analysis == "Ions") %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) 

ions_data_long_ppm = 
  ions_data %>% 
  dplyr::select(-ends_with("_flag")) %>% 
  pivot_longer(ends_with("_ppm"), names_to = "ion", values_to = "ppm") %>% 
  mutate(ion = str_remove(ion, "_ppm")) %>% 
  filter(!is.na(ppm))

ions_data_long_flag = 
  ions_data %>% 
  dplyr::select(-ends_with("_ppm")) %>% 
  pivot_longer(ends_with("_flag"), names_to = "ion", values_to = "flag") %>% 
  mutate(ion = str_remove(ion, "_flag")) %>% 
  filter(!is.na(flag))

ions_data_long_with_flags = 
  ions_data_long_ppm %>% 
  left_join(ions_data_long_flag)

ions_data_long = 
  ions_data_long_with_flags %>% 
  filter(is.na(flag))

```

```{r ions_jitter, fig.width=12, fig.height=8}
ions_data_long %>% 
  filter(ion != "nitrite") %>% 
  ggplot(aes(x = region, y = ppm))+
  geom_jitter(width = 0.1, )+
  facet_wrap(~ion, scales = "free_y")

```

```{r ions_maps}

make_map(DAT = ions_data_long %>% filter(ion == "nitrate") %>% filter(ppm <= 60), 
          VAR = "ppm", 
          COLOR_LEGEND = "Nitrate, ppm") + ggtitle("Surface waters: nitrate")

make_map(DAT = ions_data_long %>% filter(ion == "nitrate") %>% filter(ppm <= 60) %>% mutate(nitrate_n_ppm = ppm * 14/62), 
          VAR = "ppm", 
          COLOR_LEGEND = "Nitrate-N, ppm") + ggtitle("Surface waters: nitrate-N")

make_map(DAT = ions_data_long %>% filter(ion == "sulfate"), 
          VAR = "ppm", 
          COLOR_LEGEND = "Sulfate, ppm") + ggtitle("Surface waters: sulfate")

```

```{r maps_nitrite, eval=FALSE}

make_map(DAT = ions_data %>% dplyr::select(kit_id, nitrite_ppm), VAR = "nitrite_ppm", COLOR_LEGEND = "Nitrite,  ppm") + ggtitle("Surface waters: nitrite")
 
make_map(DAT = ions_data %>% dplyr::select(kit_id, nitrite_ppm) %>% 
            mutate(nitrite_n_ppm = nitrite_ppm * 14/46), 
          VAR = "nitrite_n_ppm", 
          COLOR_LEGEND = "Nitrite-N, ppm") + ggtitle("Surface waters: nitrite")
 
 
make_map(DAT = ions_data_long %>% filter(ion == "nitrite") %>% 
            mutate(nitrite_n_ppm = ppm * 14/46), 
          VAR = "nitrite_n_ppm", 
          COLOR_LEGEND = "Nitrite-N, ppm") + ggtitle("Surface waters: nitrite")
```

---

## Water: CDOM Indices

```{r cdom}
cdom_data = 
  all_data_processed %>% 
  filter(grepl("CDOM", analysis)) %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) %>% 
  dplyr::select(campaign, kit_id, transect_location, analysis, region, 
                SUVA254, Sr, FRESH, HIX, BIX, FI)

cdom_data_long = 
  cdom_data %>% 
  pivot_longer(cols = where(is.numeric)) %>% 
  filter(!is.na(value))
```

```{r cdom_jitter, fig.width=9, fig.height=6}
cdom_data_long %>% 
  ggplot(aes(x = region, y = value))+
  geom_jitter(width = 0.1, show.legend = FALSE)+
  facet_wrap(~name, scales = "free_y")
```

```{r cdom_maps}
make_map(DAT = cdom_data %>% dplyr::select(kit_id, SUVA254), VAR = "SUVA254", COLOR_LEGEND = "SUVA254") + ggtitle("Surface waters: SUVA254")
make_map(DAT = cdom_data %>% dplyr::select(kit_id, FI), VAR = "FI", COLOR_LEGEND = "FI") + ggtitle("Surface waters: fluorescence index")

#make_map(DAT = cdom_data %>% dplyr::select(kit_id, FRESH), VAR = "FRESH", COLOR_LEGEND = "FRESH")
#make_map(DAT = cdom_data %>% dplyr::select(kit_id, HIX), VAR = "HIX", COLOR_LEGEND = "HIX")

```

---

## Water: Dissolved organic carbon and total dissolved nitrogen 

DOC measured as NPOC (non-purgeable organic carbon)

```{r}
npoc_data = 
  all_data_processed %>% 
  filter(analysis == "NPOC_TDN") %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) 
```

```{r npoc_tdn_jitter, fig.width=6, fig.height=4}
npoc_data %>% 
  ggplot(aes(x = region, y = npoc_mgl))+
  geom_jitter(width = 0.1)+
  labs(y = "NPOC, mg/L")+
  #  facet_wrap(~region)+
  NULL

npoc_data %>% 
  ggplot(aes(x = region, y = tdn_mgl))+
  geom_jitter(width = 0.1)+
  labs(y = "TDN, mg/L")+
  #  facet_wrap(~region)+
  NULL
```

```{r npoc_map_old, eval=FALSE}
## This script is designed to make several figures to be used for looking at the
## data coming from EC1 with collaborators at the 2.2 Workshop in 02/22
##
## 2022-02-15 (updated 6/2/22 to be project-agnostic)
## Peter Regier 
## 
# #############
# #############

# 1. Setup ---------------------------------------------------------------------

## Load packages
require(pacman)
p_load(cowplot, ## plot_grid()
       tidyverse, 
       sf, ## simple features
       janitor, # clean_names()
       stringr,
       googlesheets4, 
       googledrive, ## import data from Google Sheets
       ggthemes) ## theme_map()

## Set CRS
common_crs <- 4326

## Set ggplot theme
# theme_set(theme_bw())

## Set map size and point size
point_size <- 2
map_width = 9
map_height = 6

# 2. Import datasets -----------------------------------------------------------
## For this example, we will use salinity which is uploaded in the water_quality
## dataset

## Read in npoc data
npoc <- npoc_data %>% 
  dplyr::select(kit_id, npoc_mgl)

latlong <- read_sheet("https://docs.google.com/spreadsheets/d/1_tKn5vHryVMHvuQix_PRXhzQcEkeAG5LDCBXjs7nZEc/edit#gid=842964887", sheet = 1) %>% 
  mutate(kit_id = paste0("K", Kit_ID)) %>% 
  rename("lat" = `Water_Latitude`, 
         "long" = `Water_Longitude`) %>% 
  select(kit_id, lat, long)

# 3. Set up map layers for plotting --------------------------------------------

## Set regional and CB (inset) bounding boxes
us_bbox <- c(xmin = -125, xmax = -60, ymin = 20, ymax = 50)
region_bbox <- c(xmin = -95, xmax = -70, ymin = 35, ymax = 48)
cb_bbox <- c(xmin = -77.8, xmax = -74.5, ymin = 36.8, ymax = 40.5)

## Make US states map cropped to GL/CB region
us <- read_sf("Reports_markdown/cb_2018_us_state_5m/cb_2018_us_state_5m.shp") %>% 
  st_transform(., crs = common_crs) %>% 
  st_crop(., y = us_bbox)

region <- st_crop(us, y = region_bbox)

## Further crop states to CB region
cb_states <- st_crop(region, y = cb_bbox)


# make a dataset merging metadata with site lat-longs
df_map <- inner_join(npoc, latlong, by = "kit_id") %>% 
  st_as_sf(., coords = c("long", "lat"), crs = common_crs)

## Crop data  to CB region for inset plot
df_cb <- st_crop(df_map, y = cb_bbox)


## Make the base map with all sites
base_plot <- ggplot() + 
  geom_sf(data = region) + 
  #geom_sf(data = meta_map, size = point_size * 3, color = "white") +
  geom_sf(data = df_map, color = "white",
          size = point_size * 2.5) + 
  geom_sf(data = df_map, aes(color = npoc_mgl), 
          size = point_size * 2) + 
  geom_rect(aes(xmin = -77.8, xmax = -74.5, ymin = 36.8, ymax = 40.5), 
            fill = NA, color = "black", lwd = 0.75) +
  geom_segment(aes(x = -77.8, xend = -76, y = 40.5, yend = 42), 
               color = "black", lwd = 0.75) + 
  geom_segment(aes(x = -74.5, xend = -71, y = 40.5, yend = 41.5), 
               color = "black", lwd = 0.75) + 
  theme_map() + 
  theme(legend.background = element_rect(fill = alpha("white", 0.0)), 
        legend.key = element_rect(fill = "transparent"), 
        legend.position = c(0.9, 0.1)) + 
  scale_color_viridis_c() + 
  labs(x = "", y = "", color = "NPOC \n (mg/L)") + 
  grattantheme::watermark("Preliminary", fontsize = 40)

## Make the inset map with just CB sites
inset_plot <- ggplot() + geom_sf(data = cb_states) + 
  geom_sf(data = df_cb, color = "white", size = point_size * 2) + 
  geom_sf(data = df_cb, aes(color = npoc_mgl), size = point_size * 1.7) + 
  theme_map() + 
  theme(legend.background = element_rect(fill = alpha("white", 0.0)), 
        legend.key = element_rect(fill = "transparent"), 
        legend.position = c(0.7, 0)) + 
  scale_color_viridis_c() + 
  theme(panel.background = element_rect(fill = "white", colour = "black")) + 
  labs(color = "")

## Combine into single figure
base_plot + 
  annotation_custom(
    ggplotGrob(inset_plot), 
    xmin = -78, xmax = -70, ymin = 41, ymax = 48.8) 
```

```{r npoc_tdn_map}
make_map(DAT = npoc_data %>% dplyr::select(kit_id, npoc_mgl), VAR = "npoc_mgl", COLOR_LEGEND = "NPOC, mg/L") + ggtitle("Surface waters: NPOC")

make_map(DAT = npoc_data %>% dplyr::select(kit_id, tdn_mgl), VAR = "tdn_mgl", COLOR_LEGEND = "TDN, mg/L") + ggtitle("Surface waters: total dissolved N")
```


---

## Water: Total suspended solids

```{r}
tss_data = 
  all_data_processed %>% 
  filter(analysis == "TSS") %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) 
```

```{r tss_jitter, fig.width=6, fig.height=4}
tss_data %>% 
  ggplot(aes(x = region, y = tss_mg_perl))+
  geom_jitter(width = 0.1)+
  labs(y = "TSS, mg/L")+
  #  facet_wrap(~region)+
  NULL

```

```{r tss_map}
make_map(DAT = tss_data %>% dplyr::select(kit_id, tss_mg_perl), VAR = "tss_mg_perl", COLOR_LEGEND = "TSS, mg/L") + ggtitle("Surface waters: TSS")
```

---

## Water: Water quality

```{r}
waterquality_data = 
  all_data_processed %>% 
  filter(analysis == "WaterQuality") %>% 
  select_if(~!(all(is.na(.)) | all(. == ""))) 


waterquality_data_long = 
  waterquality_data %>% 
  pivot_longer(cols = where(is.numeric)) %>% 
  filter(!is.na(value))
```

```{r waterquality_jitter, fig.width=12, fig.height=4}
waterquality_data_long %>% 
  ggplot(aes(x = region, y = value))+
  geom_jitter(width = 0.1)+
  facet_wrap(~name, scales = "free_y", nrow = 1)
```

```{r waterquality_map}
make_map(DAT = waterquality_data %>% dplyr::select(kit_id, ph), VAR = "ph", COLOR_LEGEND = "pH") + ggtitle("Surface waters: pH")
make_map(DAT = waterquality_data %>% dplyr::select(kit_id, sal_psu), VAR = "sal_psu", COLOR_LEGEND = "Salinity, PSU") + ggtitle("Surface waters: salinity")
```

---

---

<details>
<summary> Session Info: Click to open </summary>

Date run: `r Sys.Date()`

```{r}

sessionInfo()
```


