---
title: "submission checks"
output: 
  html_document:
    toc: true
    toc_float: true
---

#### Checks on Level 2 data before marking as "ready for submission" to ESS-DIVE

#### Created October 21, 2022 | Stephanie Pennington
#### Last ran: `r Sys.Date()`

**Checks/formatting we need to do:**

1.  Arrange by Kit ID
2.  Check file name structure
3.  Check for duplicate Kit IDs
4.  Check where each flagged value falls

```{r setup, include=FALSE}
pacman::p_load(ggplot2, readr, dplyr, tidyr, kableExtra, DT)

files <- list.files("~/Downloads/ESS-DIVE Submission/", recursive = TRUE, full.names = TRUE)
```

# Are file names in the correct structure?

`[Campaign]_[Sample Type]_[Analyte]_Level2.csv`

```{r file-naming, echo = FALSE}
basename(files) %>% 
  kable()
```

# Are there duplicate kit IDs?

```{r dupes-check, echo = FALSE, message=FALSE}
kit_check <- function(file_path) {
    df <- read_csv(file_path)
    
    if("transect_location" %in% colnames(df)) {
      df %>% 
      group_by(kit_id, transect_location) %>% 
      summarise(dupes = sum(duplicated(kit_id))) %>%  
      mutate(file = basename(file_path))
    } else {
      
      df %>% 
      group_by(kit_id) %>% 
      summarise(dupes = sum(duplicated(kit_id))) %>%  
      mutate(file = basename(file_path))
      
    }


    #tibble(file_name = basename(file_path), dupe_kitid = sum(duplicated(df$kit_id)))
}

lapply(files, kit_check) %>% bind_rows() -> dupes

datatable(dupes)
```

# Are there nas in any column?

```{r na-check, echo = FALSE, message=FALSE}


na_check <- function(file_path) {
    df <- read_csv(file_path)

    colSums(is.na(df)) -> t
    
    data.frame(as.list(t)) %>% 
      pivot_longer(everything(), names_to = "col_name", values_to = "na_count") %>% 
      mutate(file = basename(file_path), .before = col_name)

}

lapply(files, na_check) %>% bind_rows() -> na_checks

datatable(na_checks)

```

# Have the files passed checks?

```{r, echo = FALSE, message = FALSE}

dupes %>% 
  group_by(file, transect_location) %>% 
  summarise(dupe_sum = sum(dupes)) %>% 
  group_by(file) %>% 
  summarise(dupe_sum = sum(dupe_sum)) -> dupes_sum

na_checks %>% 
  group_by(file) %>% 
  summarise(na_sum = sum(na_count)) %>% 
  left_join(dupes_sum, by = "file") %>% 
  mutate('Pass/Fail' = case_when(na_sum | dupe_sum > 0 ~ "Fail",
                                 na_sum | dupe_sum == 0 ~ "Pass")) %>% 
  select(file, 'Pass/Fail', na_sum, dupe_sum) %>% 
  datatable() %>% 
  formatStyle(
  'Pass/Fail',backgroundColor = styleEqual(c("Pass", "Fail"), c('aquamarine', 'lightcoral')))

```
